SAT Competition 2020\tnoteref{title}

### 2021-01-04 paper/main.tex

\author[jku]{Nils Froleyks}
\ead{nils.froleyks@jku.at}
\author[cmu]{Marijn Heule}
\ead{marijn@cmu.edu}
\author[kit]{Markus Iser}
\ead{markus.iser@kit.edu}
\author[hiit]{Matti Järvisalo}
\ead{matti.jarvisalo@helsinki.fi}
\author[ctu]{Martin Suda} 
\ead{martin.suda@cvut.cz}

\address[jku] {
Institute for Formal Models and Verification, Johannes Kepler University\\
{\tt nils.froleyks@jku.at}\\[1em]
}

\address[cmu] {
Computer Science Department, Carnegie Mellon University\\
{\tt marijn@cmu.edu}\\[1em]
}

\address[kit] {
KIT Department of Informatics\\
{\tt markus.iser@kit.edu}\\[1em]
}

@@

\begin{abstract}
SAT Competition 2020 stands in the tradition of the series of annual competitive events which motivate and assess the progress in SAT solving. 
This competition was special as it introduced the new cloud track where SAT Solvers that run on hundreds of processors could compete. 
Another novelty was the application-specific sub-track of the main track, 
where solvers competed in solving benchmark instances from one specific domain only, and in this year that was the planning domain. 
We used new tools to select and distribute benchmark instances and their attributes. 
In this paper we provide a description of the well-known and the new competition tracks and how we organized them. 
It follows then a detailed analysis of the results and the strategies of the award winning solvers. 
\end{abstract}

\begin{keyword}
SAT, Competition
\end{keyword}

@@

@@

@@

By filtering out instances with could be solved by Minisat in less than $10$ minutes, we obtained an initial set of $1012$ instances. 
In order to compile a balanced set of $300$ benchmark instances, we considered a maximium of $14$ submissions per author and used the procedure which is depicted in Algorithm~\ref{algo:select}. 
Per author, if possible, we first randomly selected $7$ satisfiable and $7$ unsatisfiable instances (lines~3 and~4). 
If this did not yield a total of $14$ instances, we added instances of yet unknown outcome (lines~5--7). 
Of the such obtained $308$ instances, we randomly removed $8$ satisfiable instances, yielding a total of $114$ satisfiable, $78$ unsatisfiable and $108$ instances of unknown result.

To obtain the final compilation of $400$ benchmark instances, we augmented this set with $100$ instances which have been used in previous competitions. 
We randomly selected $21$ satisfiable, $57$ unsatisfiable and $22$ unknown instances to yield a total of $135$ satisfiable, $135$ unsatisfiable and $130$ instances of unkown result. 
Furthermore, we made sure not to select instances of a family which is already represented in the set of newly submitted instances and excluded random, agile and planning instances (due to the planning track).

@@

The last big cluster in the center of the heat map is comprised of the solvers
originally written by Biere. The three configurations of \emph{Kissat} form a
subcluster, even though the configurations are fairly distinct from each other.
% , specialized for satisfiable and unsatisfiable instances.
% One targets satisfiable instances, while the other is focused on unsatisfiable
% instances and a third alternating between them.

The other subcluster consists of
modifications to the 2019 version of \emph{\mbox{CaDiCaL}}. The new version
CaDiCaL-2020 has a worse overall performance than
% has an overall worse per
% is lacking behind
the other solvers in this cluster.

@@

Another more recent string of development is rooted in the \solver{CoMinisatPS} fork of \solver{Minisat} by Chanseok Oh introducing three-tier clause-managment~\cite{Oh:2015:satunsat}. 
The SAT solver \solver{Maple} appeared as a series of forks presenting innovative branching heurstics at SAT Competition 2016~\cite{Liang:2016:LRB}. 
The then award winning variant \solver{MapleCOMSPS} is an implementation of a hybrid branching heuristic of classic Variable-State Independend Decaying Sum (VSIDS) and the new Learning-Rate based Branching (LRB) in \solver{CoMinisatPS}~\cite{Liang:2016:MapleCOMSPS}.

@@

Kochemazov et al. improved three-tier clause-management by persting additional clauses through hash-based detection of repeatedly learned clauses and presented their award winning~\solver{MapleLCMDistChronoBT-DL} to SAT Race~2019~\cite{Kochemazov:SC2019:MapleChronoBTDL}. 
As can be seen in Table~\ref{tab:solvers-incremental}, numerous submissions to SAT Competition~2020 are forks of some recently award winning \solver{Maple} solver.

@@

Many independent and award-winning code-bases can be found among the state-of-the-art SAT solvers written by Biere. 
The sequential SAT solver \solver{Lingeling} has been award winning since SAT Competition~2011 and is still competitive in its parallel version \solver{Plingeling}~\cite{Biere:2012:Lingeling}. 
As of SAT Competition~2017, \solver{CaDiCaL} by Biere is another independent award winning and structured implementation of the state-of-the-art in SAT solving. 
Its improved reimplementation \solver{Kissat} won in several of this competitions sequential tracks by a large margin~\cite{Biere:SC2020}.

@@

Sequential SAT solvers have been evaluated in the Main tracks as well as in the Incremental track of SAT Competition~2020. 
A total of $18$ teams submitted correct solvers to the Main track of the competition. 
Table~\ref{tab:solvers-main} displays an overview of the participating teams, base solvers and their variants. 
The recent genealogy of all submissions, is highlighted by \emph{base solver} and \emph{variant} columns. 
Underlined text is used to point out the important parts of the respective submissions name.

@@

\solver{Kissat} uses a new unit (``ticks'') to measure the length of two alternating restart-modes based on the approximate number of cache-line accesses, after conflict-rate has been obeserved to be too unstable for the use-case~\cite{Biere:SC2020}.

@@

The \solver{Relaxed} fork integrates short runs of the local search solver \solver{CCAnr} through periodic ex- and import of assignments~\cite{Xindi:SC2019}. 
The submitted versions of \solver{Relaxed Maple} use a probabilistic schedule for switching between $10$ phase selection modes. 
The award winning variant \solver{newTech} uses occurrence counts of variables in unsatisfied clauses during SLS runs to recalculate variable priorities for their modified branching heuristic~\cite{Xindi:SC2020}.

@@

\solver{F2TRC} introduces an improved management of learned clauses in the tree tiers \emph{core}, \emph{tier2} and \emph{local} which are inherited from \solver{CoMinisatPS}~\cite{Oh:2015:satunsat}.
A dynamic size limit for the \emph{core} tier, triggers the reassignment of inactive clauses from \emph{core} to \emph{tier2}. 
To conter-act an observered starving of tier2, the conflict-based heuristic that controls demotion of clauses from \emph{tier2} to \emph{local} was replaced by a size based heuristic~\cite{Kochemazov:SC2020}.

@@

\solver{Painless-MCOMSPS-STR} integrates the solver \solver{MapleCOMSPS} in the \solver{Painless} parallelization framework~\cite{Frioux:2017:Painless}~\cite{Liang:2017:Maplecomsps}. 
The authors submitted a \solver{32 core} and a \solver{64 core} variant and were awared three prices: first place in the Main track, first place in the Sat track and second in the Unsat track. 
The \solver{32 core} variant achieved slightly better scores than the \solver{64 core} variant. 
The differentiated evaluation of the Sat and Unsat tracks shows the \solver{32 core} variant performed much better, such that the follow-up solvers outperformed the \solver{64 core} variant.

\solver{Painless} uses a generic interface to integrate a solver and abstracts away the implementation details of parallelism and concurrenct data-structures, such that implementations in \solver{Painless} boil down to implementing parallelization and clause sharing strategies.

\solver{Painless-MCOMSPS-STR} diversifies mainly via diverse fixed configurations of the branching strategies LRB and VSIDS, and sparse random intialization of variable polarities~\cite{Balyo:2015:Hordesat}. 
However, a special solver instance performs concurrent clause strengthening~\cite{Wieringa:2013:CCS}. 
Another special solver instance is configured to perform Gaussian eliminiation during preprocessing. 
\solver{Painless-MCOMSPS-STR} uses an all-to-all strategy for clause sharing with a fixed size clause buffer and an LBD filter with a dynamic limit.

@@

\solver{ManyGlucose} was submitted in \solver{32 core} and \solver{64 core} variants. 
The \solver{32 core} variant received two awards in this competition: one third place in the Main track and another third place in the Unsat track. 
\solver{ManyGlucose} is a fork of \solver{GlucoseSyrup} that uses strategies known from \solver{ManySat} to achive deterministic solver behaviour~\cite{GlucoseSyrup}~\cite{ManySat}.

@@

\solver{Painless Maple} received the award for second best performance in the Sat track but scored last in the Unsat track. 
\solver{Painless Maple} integrates the solver \solver{ExMapleLCMDistChronoBT}, which is a modified \solver{MapleLCMDistChronoBT} (see Main track), in the \solver{Painless} parallelization framework.

@@

\solver{Mallob} shares clauses by organizing solvers in a binary tree in which clauses are asynchronously aggregated in a buffer which is passed along this tree from leafs to the root. 
Each node performs a three-way merge of its local export and the two incoming buffers. 
The aggragate that approaches the root of the binary tree is then broadcast to all solvers~\cite{}.

@@

\solver{TopoSAT~2} is a massively parallel SAT solver build from \solver{Glucose~3}~\cite{Glucose3,Toposat2}. 
\solver{TopoSAT~2} uses lockfree clause-exchange for solvers running on the same maching, and MPI to share clauses between machines~\cite{ManySat,Proceedings}.

@@

\solver{Slime} is build from \solver{MapleLCMDistChronoBT} and was first submitted as a sequential solver to SAT Race~2019 with a new phase selection heuristic~\cite{SlimeSATRace2019}. 
The new submitted version of \solver{Slime} comes with periodic randomization in geometrically increasing intervals~\cite{}. 
Even thought its sequential version was unsuccessful in the Main track, the MPI-based cloud version of \solver{Slime} won the third price in the Cloud track.

@@

\section*{Acknowledgements}
Martin Suda was supported by the ERC Consolidator grant AI4REASON no. 649043 under the EU-H2020 programme,
the Czech Science Foundataion project 20-06390Y, and the project RICAIP, no. 857306 under the EU-H2020 programme.

>>>>

\author[jku]{Nils Froleyks}
\ead{nils.froleyks@jku.at}
\author[cmu]{Marijn Heule}
\ead{marijn@cmu.edu}
\author[kit]{Markus Iser}
\ead{markus.iser@kit.edu}
\author[uh]{Matti J\"arvisalo}
\ead{matti.jarvisalo@helsinki.fi}
\author[ctu]{Martin Suda} 
\ead{martin.suda@cvut.cz}

\address[jku] {
Institute for Formal Models and Verification, Johannes Kepler University, Austria\\
{\tt nils.froleyks@jku.at}\\[1em]
}

\address[cmu] {
Computer Science Department, Carnegie Mellon University, USA\\
{\tt marijn@cmu.edu}\\[1em]
}

\address[kit] {
Department of Informatics, Karlsruhe Institute of Technology, Germany \\
{\tt markus.iser@kit.edu}\\[1em]
}

\address[uh] {
HIIT, Department of Computer Science, University of Helsinki, Finland
{\tt matti.jarvisalo@helsinki.fi}\\[1em]
}

@@

\begin{abstract}
The SAT Competitions constitute a well-established series of yearly open international competitive algorithm implementation competitions
focusing on the Boolean satisfiability (or propositional satisfiability, SAT) problem. 
In this article, we provide
a detailed account on the 2020 instantiation of the SAT Competition, including the 
 new competition tracks and benchmark selection procedures, overview of solving strategies implemented in top-performing solvers, 
and a more detailed analysis of the empirical data obtained from running the competition.
%SAT Competition 2020 stands in the tradition of the series of annual competitive events which motivate and assess the progress in SAT solving. 
%This competition was special as it introduced the new cloud track where SAT Solvers that run on hundreds of processors could compete. 
%Another novelty was the application-specific sub-track of the main track, 
%where solvers competed in solving benchmark instances from one specific domain only, and in this year that was the planning domain. 
%We used new tools to select and distribute benchmark instances and their attributes. 
%In this paper we provide a description of the well-known and the new competition tracks and how we organized them. 
%It follows then a detailed analysis of the results and the strategies of the award winning solvers. 
\end{abstract}

\begin{keyword}
SAT, Boolean satisfiability, SAT Competition, SAT solvers, empirical evaluation,
benchmarking
\end{keyword}

@@

From what was once mainly the archetypal intractable (in particular NP-complete) problem, propositional satisfiability 
(or Boolean satisfiability, SAT) has flourished in practice into a success story of modern computer science.
This is due to advances in SAT solvers, i.e., implementations of decision procedures for SAT, which form today a 
central computational tool for solving real-world problem instances of various kinds of NP-hard search and optimization problems.
With standardized inputformats, readily-available APIs for incrementalapplications,
and certified proof logging andchecking capabilities, applications of SAT solver technology
have branched from first breakthrough applications in automated planning, test pattern generation and hardware verification
to thousands of different application settings.

The success of SAT would not be without the persistent efforts of the SAT community to further develop the performance and robustness
of SAT solvers. The SAT Competition series, which a history dating back to early 90s, aims to support and provide further incentives 
for maintaining and furthering this progress. In particular,

WILL CONTINUE FROM HERE -MATTI

===

OLD:

@@

OLD END ===

@@

By filtering out instances with could be solved by Minisat in less than $10$ minutes, we obtained an initial set of $1012$ instances. 
In order to compile a balanced set of $300$ benchmark instances, we considered a maximum of $14$ submissions per author and used the procedure which is depicted in Algorithm~\ref{algo:select}. 
Per author, if possible, we first randomly selected $7$ satisfiable and $7$ unsatisfiable instances (lines~3 and~4). 
If this did not yield a total of $14$ instances, we added instances of yet unknown outcome (lines~5--7). 
Of the such obtained $308$ instances, we randomly removed $8$ satisfiable instances, yielding a total of $114$ satisfiable, $78$ unsatisfiable and $108$ instances of unknown result.

To obtain the final compilation of $400$ benchmark instances, we augmented this set with $100$ instances which have been used in previous competitions. 
We randomly selected $21$ satisfiable, $57$ unsatisfiable and $22$ unknown instances to yield a total of $135$ satisfiable, $135$ unsatisfiable and $130$ instances of unknown result. 
Furthermore, we made sure not to select instances of a family which is already represented in the set of newly submitted instances and excluded random, agile and planning instances (due to the planning track).

@@

The last big cluster in the center of the heat map is comprised of the solvers
originally written by Biere. The three configurations of \emph{Kissat} form a
sub-cluster, even though the configurations are fairly distinct from each other.
% , specialized for satisfiable and unsatisfiable instances.
% One targets satisfiable instances, while the other is focused on unsatisfiable
% instances and a third alternating between them.

The other sub-cluster consists of
modifications to the 2019 version of \emph{\mbox{CaDiCaL}}. The new version
CaDiCaL-2020 has a worse overall performance than
% has an overall worse per
% is lacking behind
the other solvers in this cluster.

@@

Another more recent string of development is rooted in the \solver{CoMinisatPS} fork of \solver{Minisat} by Chanseok Oh introducing three-tier clause-management~\cite{Oh:2015:satunsat}. 
The SAT solver \solver{Maple} appeared as a series of forks presenting innovative branching heuristics at SAT Competition 2016~\cite{Liang:2016:LRB}. 
The then award winning variant \solver{MapleCOMSPS} is an implementation of a hybrid branching heuristic of classic Variable-State Independent Decaying Sum (VSIDS) and the new Learning-Rate based Branching (LRB) in \solver{CoMinisatPS}~\cite{Liang:2016:MapleCOMSPS}.

@@

Kochemazov et al. improved three-tier clause-management by persting \todo{Please check "persting"!}
additional clauses through hash-based detection of repeatedly learned clauses and presented their award winning~\solver{MapleLCMDistChronoBT-DL} to SAT Race~2019~\cite{Kochemazov:SC2019:MapleChronoBTDL}. 
As can be seen in Table~\ref{tab:solvers-incremental}, numerous submissions to SAT Competition~2020 are forks of some recently award winning \solver{Maple} solver.

@@

Many independent and award-winning code-bases can be found among the state-of-the-art SAT solvers written by Biere. 
The sequential SAT solver \solver{Lingeling} has been award winning since SAT Competition~2011 and is still competitive in its parallel version \solver{Plingeling}~\cite{Biere:2012:Lingeling}. 
As of SAT Competition~2017, \solver{CaDiCaL} % \cite{Biere:SC2020} - or cite some older proceedings to match the "As of SAT Competition~2017" part?
by Biere is another independent award winning and structured implementation of the state-of-the-art in SAT solving. 
Its improved reimplementation \solver{Kissat}~\cite{Biere:SC2020} won in several of this competitions sequential tracks by a large margin.

@@

Sequential SAT solvers have been evaluated in the Main tracks as well as in the Incremental track of SAT Competition~2020. 
A total of $18$ teams submitted correct solvers to the Main track of the competition. 
Table~\ref{tab:solvers-main} displays an overview of the participating teams, base solvers and their variants. 
The recent genealogy of all submissions is highlighted by \emph{base solver} and \emph{variant} columns. 
Underlined text is used to point out the important parts of the respective submissions name.

@@

\solver{Kissat} uses a new unit (``ticks'') to measure the length of two alternating restart-modes based on the approximate number of cache-line accesses, after conflict-rate has been observed to be too unstable for the use-case~\cite{Biere:SC2020}.

@@

The \solver{Relaxed} fork integrates short runs of the local search solver \solver{CCAnr} through periodic export and import of assignments~\cite{Xindi:SC2019}. 
The submitted versions of \solver{Relaxed Maple} use a probabilistic schedule for switching between $10$ phase selection modes. 
The award winning variant \solver{newTech} uses occurrence counts of variables in unsatisfied clauses during SLS runs to recalculate variable priorities for their modified branching heuristic~\cite{Xindi:SC2020}.

@@

\solver{F2TRC} introduces an improved management of learned clauses in the tree tiers \emph{core}, \emph{tier2} and \emph{local} which are inherited from \solver{CoMinisatPS}~\cite{Oh:2015:satunsat}.
A dynamic size limit for the \emph{core} tier triggers the reassignment of inactive clauses from \emph{core} to \emph{tier2}. 
To counter-act an observed starving of tier2, the conflict-based heuristic that controls demotion of clauses from \emph{tier2} to \emph{local} was replaced by a size based heuristic~\cite{Kochemazov:SC2020}.

@@

\solver{Painless-MCOMSPS-STR} integrates the solver \solver{MapleCOMSPS} in the \solver{Painless} parallelization framework~\cite{Frioux:2017:Painless,Liang:2017:Maplecomsps}. 
The authors submitted a \solver{32 core} and a \solver{64 core} variant and were awarded three prices: first place in the Main track, first place in the Sat track and second in the Unsat track. 
The \solver{32 core} variant achieved slightly better scores than the \solver{64 core} variant. 
The differentiated evaluation of the Sat and Unsat tracks shows the \solver{32 core} variant performed much better, such that the follow-up solvers outperformed the \solver{64 core} variant.

\solver{Painless} uses a generic interface to integrate a solver and abstracts away the implementation details of parallelism and concurrent data-structures, such that implementations in \solver{Painless} boil down to implementing parallelization and clause sharing strategies.

\solver{Painless-MCOMSPS-STR} diversifies mainly via diverse fixed configurations of the branching strategies LRB and VSIDS, and sparse random initialization of variable polarities~\cite{Balyo:2015:Hordesat}. 
However, a special solver instance performs concurrent clause strengthening~\cite{Wieringa:2013:CCS}. 
Another special solver instance is configured to perform Gaussian elimination during preprocessing. 
\solver{Painless-MCOMSPS-STR} uses an all-to-all strategy for clause sharing with a fixed size clause buffer and an LBD filter with a dynamic limit.

@@

\solver{ManyGlucose} was submitted in \solver{32 core} and \solver{64 core} variants. 
The \solver{32 core} variant received two awards in this competition: one third place in the Main track and another third place in the Unsat track. 
\solver{ManyGlucose} is a fork of \solver{GlucoseSyrup} that uses strategies known from \solver{ManySat} to achieve deterministic solver behaviour~\cite{GlucoseSyrup}~\cite{ManySat}.

@@

\solver{Painless Maple} received the award for the second best performance in the Sat track but scored last in the Unsat track. 
\solver{Painless Maple} integrates the solver \solver{ExMapleLCMDistChronoBT}, which is a modified \solver{MapleLCMDistChronoBT} (see Main track), in the \solver{Painless} parallelization framework.

@@

\solver{Mallob} shares clauses by organizing solvers in a binary tree in which clauses are asynchronously aggregated in a buffer which is passed along this tree from leafs to the root. 
Each node performs a three-way merge of its local export and the two incoming buffers. 
The aggregate that approaches the root of the binary tree is then broadcast to all solvers~\cite{}.

@@

\solver{TopoSAT~2} is a massively parallel SAT solver build from \solver{Glucose~3}~\cite{Glucose3,Toposat2}. 
\solver{TopoSAT~2} uses lock-free clause-exchange for solvers running on the same matching, and MPI \todo{Please spell out MPI once.}
to share clauses between machines~\cite{ManySat,Proceedings}.

@@

\solver{Slime} is built from \solver{MapleLCMDistChronoBT} and was first submitted as a sequential solver to SAT Race~2019 with a new phase selection heuristic~\cite{SlimeSATRace2019}. 
The new submitted version of \solver{Slime} comes with periodic randomization in geometrically increasing intervals~\cite{}. 
Even thought its sequential version was unsuccessful in the Main track, the MPI-based cloud version of \solver{Slime} won the third price in the Cloud track.

@@

\section*{Acknowledgements}
Matti J\"arvisalo was financially supported by Academy of Finland grants 322869 and 328718.
Martin Suda was supported by the ERC Consolidator grant AI4REASON no. 649043 under the EU-H2020 programme,
the Czech Science Foundation project 20-06390Y, and the project RICAIP, no. 857306 under the EU-H2020 programme.
